Evaluating Informative Hypotheses with Equality and Inequality Constraints:\\ A tutorial using the Bayes Factor via the Encompassing Prior Approach

### 2021-11-10 Documents/Paper/paper.Rnw
update revision

\title{Evaluating Informative Hypotheses with Equality and Inequality Constraints: The Bayes Factor via the Encompassing Prior Approach}
\shorttitle{The Bayes Factor via the Encompassing Prior Approach}

@@

\abstract{When conducting a study, researchers usually have expectations based on hypotheses or theoretical perspectives they want to evaluate. Equality and inequality constraints on the model parameters are used to formalize researchers' expectations or theoretical perspectives into the so-called “\textit{informative hypotheses}”. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are unsuitable for testing complex informative hypotheses. An alternative approach is to use the Bayes Factor. In particular, the Bayes Factor based on the “\textit{encompassing prior approach}” allows researchers to easily evaluate complex informative hypotheses. In this paper, we provide a clear and detailed introduction to this approach. We describe all the steps and elements involved in the formalization of informative hypotheses and computation of the Bayes Factor following the encompassing prior approach. Subsequently, we provide an application of this method to a real case scenario comparing the different theoretical perspectives regarding the role of mother attachment and father attachment on children's social-emotional development.
\vspace{2em}
\begin{center}
\textbf{Translational Abstract}
\end{center}
\vspace{.5em}
\noindent The evaluation of hypotheses is one of the main goals of empirical research. In fact, when conducting a study, researchers usually want to evaluate the plausibility of specific hypotheses according to the collected data. These hypotheses can be formalized in accordance with researchers' expectations or theoretical perspectives into the so-called “\textit{informative hypotheses}”. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are not suitable for evaluating the plausibility of complex informative hypotheses. The Bayes Factor via the “\textit{encompassing prior approach}”, instead, allows researchers to easily evaluate the plausibility of complex informative hypotheses. In this paper, we provide a clear and detailed introduction to this approach. We describe all the steps and elements involved in the formalization of informative hypotheses and computation of the Bayes Factor. Moreover, we provide an application of this method to a real case scenario comparing the different theoretical perspectives regarding the role of mother attachment and father attachment on children's social-emotional development. This approach is of particular interest in applied research as it allows researchers to formalize hypotheses with great flexibility and to directly evaluate which hypothesis is the most supported by the data.
}

\keywords{Bayes Factor, Encompassing Prior, Informative Hypothesis, Equality and Inequality Constraints, Attahcement Theory}

\authornote{
   \addORCIDlink{Claudio Zandonella Callegher}{0000-0001-7721-6318}

  Correspondence concerning this article should be addressed to Claudio Zandonella Callegher, Department of Developmental Psychology and Socialisation, University of Padua, Via Venezia, 8 - 35131 Padova, Italy.  E-mail: claudiozandonella@gmail.com}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

@@

In psychology, the dominant statistical approach to evaluate research hypotheses is the Null Hypothesis Significance Testing (NHST). In the literature, however, the utility and validity of NHST are largely debated \parencite{wassersteinMovingWorld052019}. This approach presents indeed several limitations. First, the NHST places a narrow focus on statistical testing rather than on the formalization of hypotheses. This has usually led researchers to evaluate data against the catch-all null-hypothesis that nothing is going on rather than testing their specific expectation [and the alternative hypothesis is rarely formalized]. Second, the p-value does not quantify the evidence in favour of one hypothesis, thus, it is not possible to \textit{“accept”} a hypothesis but only to \textit{“reject”} it. This is inconvenient as, in case of not significant results, the researchers are left in a state of indecision. Third, NHST does not allow testing multiple hypotheses at the same time. Using the NHST the null hypothesis is tested only against a single alternative. Fourth, NHST is unsuitable for testing broad classes of hypotheses with equality and inequality constraints \parencite{mulderSimpleBayesianTesting2019}. Expectations can be evaluated using one-side tests, however, when more groups or more variables are involved, it is not possible to evaluate complex parameter constraints that reflect researchers' expectations \parencite{vandeschootIntroductionBayesianModel2011}.

Model comparison is a different approach that allows researchers to compare multiple hypotheses and identify which is the most supported by the data \parencite{mcelreathStatisticalRethinkingBayesian2020}. Hypotheses are first formalized as statistical models according to researchers expectations or theoretical perspectives. Subsequently, it is possible to evaluate which is the most supported model among those considered according to the data. To do that a popular approach is to use information criteria such as the AIC or BIC criteria that estimate models ability to predict new data \parencite{wagenmakersAICModelSelection2004,akaike1973a, schwarzEstimatingDimensionModel1978}. Model comparison has several advantages and, in particular, it allows to directly compute the relative plausibility of each model given the data and the set of models considered. Model comparison using the information criteria, however, is not appropriate to evaluate hypotheses that include complex parameter constraints reflecting researchers' expectations. In fact, information criteria evaluate models complexity according to the number of parameters, but they do not take into account possible order constraints on the parameters.

An alternative criterium to evaluate research hypotheses in a model comparison is the Bayes Factor. In the last 25 years, there has been an increasing interest in the Bayes Factor and its use has been proposed as the solution to the critical issues of the NHST \parencite{heckReviewApplicationsBayes2020, mulderEditorsIntroductionSpecial2016}. Although it has its own limitations and it does not solve the fundamental issues of the misuse of statistical techniques \parencite{schadWorkflowTechniquesRobust2021, gelmanBayesianDataAnalysis2013}, the Bayes Factor offers some clear advantages. In particular, the Bayes Factor allows us to compare hypotheses obtaining a relative index of evidence, like the information criteria, but it also allows us to easily compare complex research hypotheses. In fact, so-called \textit{“informative hypotheses”} (i.e., hypotheses containing information about the ordering of the model parameters) can be formalized according to researchers' expectations or theoretical perspectives using equality and inequality constraints \parencite{vandeschootIntroductionBayesianModel2011}. Subsequently, these hypotheses can be compared against each other using the Bayes Factor.

The evaluation of informative hypotheses is of particular interest to the researchers as it allows them to directly assess specific complex expectations and theoretical perspectives. In the literature, a particular approach allowing to easily compute the Bayes Factor with complex informative hypotheses has received increasing attention: \textit{“the Bayes Factor with encompassing prior”}. \textcite{vandeschootIntroductionBayesianModel2011} presented a general introduction to informative hypotheses testing using the Bayes Factor with the encompassing prior approach, whereas \textcite{hoijtinkInformativeHypothesesTheory2012} offered a more detailed description of the development of this method. Other studies, instead, considered the application of this approach with specific statistical models; for example: mixed effect models \parencite{katoBayesianApproachInequality2006}, evaluation of correlation coefficients \parencite{mulderBayesFactorsTesting2016} and multinomial models \parencite{heckMultinomialModelsLinear2019}. In addition, \textcite{guBayesianEvaluationInequality2014} proposed a general approximate procedure to evaluate inequality constraints in a wide range of statistical models.

All studies mentioned above, however, considered only informative hypotheses with inequality constraints. Evaluating in the same hypothesis equality and inequality constraints, instead, it was no possible and equality constraints had to be approximate to “\textit{about equality constraints}” (i.e., a equality constraints of type $\theta_i = \theta_j$ is approximated to $|\theta_i - \theta_j| < \xi$ for a value of $\xi$ small enough).

@@

The development and implementation of this approach are of particular interest because, although with its limits, it allows researchers to directly evaluate their expectations and hypotheses. The available literature, however, is rather technical. The complexity of these articles makes it difficult for researchers not familiar with this approach to clearly understand all the steps involved. On the other hand, articles offering a general introduction to the Bayes Factor do not provide enough details to allow readers to autonomously apply this method to their problems, but they usually rely on ad-hoc solutions implemented in some statistical software.

@@

Prior specification is an important element as the resulting Bayes Factor value is affected by the prior choice. This aspect is particularly relevant in the case of equality constraints \parencite{lindleyStatisticalParadox1957, bartlettCommentLindleyStatistical1957}, whereas inequality constraints are not affected as long as the prior is symmetric and centred to the focal point of interest (this aspect will be further discussed in Section~\ref{sec:adj-prior}).

@@

Both, \textcite{guApproximatedAdjustedFractional2018, mulderSimpleBayesianTesting2019} based their approach on the fractional Bayes Factor. Starting from a non-informative prior, a minimal fraction of the data is used to obtain a posterior that is subsequently used as proper prior, we refer to it as \textit{fractional prior} (have you ever heard Lindley's \citeyear{lindleyBayesianStatisticsReview1972} famous quote \textit{“Today’s posterior is tomorrow’s prior”}?). The remaining part of the data is used to compute the Bayes Factor. The two approaches, however, have an important difference. \textcite{guApproximatedAdjustedFractional2018} approximated the obtained fractional prior (as well as the posterior, see Section~\ref{sec:posterior-encompassing}) to a (multivariate) normal distribution. In fact, due to large-sample theory \parencite{gelmanBayesianDataAnalysis2013}, parameter posterior distribution can be approximated to a (multivariate) normal distribution. On the contrary, \textcite{mulderSimpleBayesianTesting2019} provided an analytic solution in the case of linear regression models, obtaining an accurate quantification of the distribution.

All methods discussed above suggest objective procedures to avoid arbitrary prior specification. Nevertheless, using subjective (reasonable) priors according to previous information or experts' indications is still a possible approach. Keep in mind, however, that prior specification (even if obtained through an \textit{objective procedure}) affects the Bayes Factor results. Thus, it is fundamental to conduct a prior sensitivity analysis to evaluate the influence of prior specification on the results \parencite{schadWorkflowTechniquesRobust2021, duBayesFactorOnesample2019}.

@@

At this point, several authors underline the importance of centring the prior distribution on the constraints points of interest. Let's further discuss this issue in the next section.

@@

At this point, we have both the prior and the posterior distributions of the parameters of interests of the encompassing model. Before proceeding, however, let's underline some important aspects.

@@

Analogously, it is possible to compute the \textbf{marginal posterior density} $\pi_{adj}(\beta_E =0| Y, H_u)$ considering this time the estimated posterior mean vector $\bm{\hat{\beta}}$ and the estimated  posterior covariance matrix $\hat{\Sigma}_{\beta}$.

@@

To compute the Bayes Factor between different informative hypotheses, we can simply consider the ratio of the Bayes Factor of each hypothesis against the encompassing model:
\begin{equation}
BF_{ij} = \frac{BF_{iu}}{BF_{ju}}
\end{equation}
In order to do that, however, it is necessary for the informative hypotheses considered to be comparable \parencite{guApproximatedAdjustedFractional2018}. That is, the intersection of the constrained region of each hypothesis must be non-empty. Thus, a set of informative hypotheses $\{H_1,\ldots, H_i\}$ is comparable if exist at least one solution to
\begin{equation}
\begin{bmatrix}
R_{1E}\\
R_{1I}
\end{bmatrix} \theta = \begin{bmatrix}
r_{1E}\\
r_{1I}
\end{bmatrix}, \ldots,\begin{bmatrix}
R_{iE}\\
R_{iI}
\end{bmatrix} \theta = \begin{bmatrix}
r_{iE}\\
r_{iI}
\end{bmatrix}.
\end{equation}

@@

The described approach using (multivariate) normal approximation can be straightforward applied when parameters of interest are unbounded. In the case of bounded parameters (i.e., variances, correlations, or probabilities), however, appropriate prior distributions are required to take into account the parameters bounded nature.

@@

Note that these issues are relevant only when considering the prior distribution. In the case of the posterior distribution, instead, we can safely approximate it to a (multivariate) normal distribution thanks to the large-sample theory. Moreover, prior specification is relevant only for the parameters of interest (i.e., the parameters involved in the constraints) as nuisance parameters are integrated out in the Bayes Factor. Thus, the results will be insensitive to the actual choice of the prior for the nuisance parameters as long as it is vague enough.

@@

\section{A Case Study: Hypotheses Testing in the Attachment Theory}

In this section, we propose a real case study evaluating different informative hypotheses within the attachment theory. This will allow us to face the common issues found in real research applications. First, we provide the required background information regarding the attachment theory and the study aims and characteristics. Subsequently, all the steps involved in the evaluation of informative hypotheses using the Bayes Factor with the encompassing prior approach are described. Finally, the results are briefly discussed considering the prior sensitivity analysis and limits of this approach.

@@

Attachment theory originates from the pioneering work of \textcite{bowlbyAttachmentLoss1969} and Ainsworth (\citeyear{ainsworthAttachmentExplorationSeparation1970}). They postulates that children in stressful situations actively seek proximity of the caregiver through some attachment behaviors (e.g., crying; moving towards the caregiver) in order to fulfil the evolutionary goal of protection from dangers. The main tenet of attachment theory is that the relationships with the caregivers that children develop in the early stages of their life (i.e., \textit{attachment bond}) will affect children’s social-emotional future development \parencite{cassidyHandbookAttachmentTheory2016}. Besides behavior, people construct mental representations, or working models, of the self and significant others based on their interpersonal experiences. Four main attachment styles have been recognized in the literature according to different internal representations:
\begin{itemize}
  \item{\textbf{Secure Attachment}} - children who are securely attached display optimal emotional regulation and they consider the caregiver as a secure base.
  \item{\textbf{Anxious Attachment}} - anxious children manifest high levels of anxiety in stressful situations and their relationships with the caregivers is ambivalent displaying anger or helplessness.
  \item{\textbf{Avoidant Attachment}} - avoidant children mask distress in stressful situations displaying little emotions and their relationships with the caregivers is characterized by little involvement.
  \item{\textbf{Fearful Attachment}} - fearful children lack adequate emotional regulation in stressful situations with the risk of displaying non organized behaviors.
\end{itemize}

The attachment theory is one of the main and most supported theories in psychology \parencite{cassidyHandbookAttachmentTheory2016}. In the literature, however, there is still an open debate on the relative role of mother and father attachment on children's social-emotional development. Four different main theoretical perspectives have been identified \parencite{brethertonFathersAttachmentTheory2010}:
\begin{itemize}
  \item{\textbf{Monotropy Theory}} - only the principal attachment figure (usually the mother) has an impact on children's development.
  \item{\textbf{Hierarchy Theory}} - the principal attachment figure (usually the mother) has a greater impact on children development than the subsidiary attachment figure (usually the father).
  \item{\textbf{Independence Theory}} - all attachment figures are equally important but they affect the children's development differently.
  \item{\textbf{Integration Theory}} - to understand the impact on children's development it is necessary to consider all attachment relationships taken together.
\end{itemize}
Contrasting results have been reported by studies investigating which is the “\textit{correct}” theory. No study, however, has tried to properly evaluate the different theoretical perspectives by directly comparing the different hypotheses. [TODO add some references]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Present Study}

@@

Children's social-emotional development was measured using the Strength \& Difficulties Questionnaire \parencite[SDQ;][]{goodmanWhenUseBroader2010} completed by the teachers. Separate scores for externalizing and internalizing problems were obtained as sum of the questionnaire items. In the analysis, however, only externalizing problems  are considered as teachers are expected to be better at reporting externalizing problems than internalizing problems [TODO spiegare meglio il perchè? aggiungere ref?]. The distribution of externalizing problems ($\text{Mean} = \Sexpr{my_round(mean(data_cluster$externalizing_sum), 2)};\ \text{SD} =  \Sexpr{my_round(sd(data_cluster$externalizing_sum), 2)};\ \text{Median} = \Sexpr{my_round(median(data_cluster$externalizing_sum), 1)}$) is presented in Figure~\ref{fig:plot-externalizing-dist}.
<<plot-externalizing-dist, cache=TRUE, fig.cap="Distribution of externalizing problems ($n_{subj} = 847$).">>=
plot_externalizing_dist()
@

Externalizing problems according to attachment styles are reported in Table~\ref{tab:table-cluster-ext}. More information about the sample, descriptive statistics, cluster analysis, and analysis of internalizing problems can be found in the Supplemental Material available online \url{https://claudiozandonella.github.io/Attachment/}.

@@

When interpreting the results, it is important to take into account the limits of the Bayes Factor. First of all, the selected hypothesis is relative to the data and the set of hypotheses considered. That means that we should not interpret the selected hypothesis as the only “\textit{correct}” one. Maybe we did not consider some important aspects and new hypotheses may reveal themselves to be actually much better than the previous ones. For example, regarding the attachment results, taking into account the interaction between children's gender and attachment may be fundamental (do you expect that attachment towards the father plays a different role between girls and boys?). This limit, however, becomes an advantage as it forces the researchers to focus on the definition of the hypotheses: to obtain good answers first we need to ask the right questions.

@@

As underlined by \textcite[p.229]{guApproximatedAdjustedFractional2018}, “this class of informative hypotheses covers a much broader range of scientific expectations than the class of standard null hypotheses. In addition, by testing competing informative hypotheses directly against each other a researcher obtains a direct answer as to which scientific theory is most supported by the data”.

>>>>

\title{Evaluating Informative Hypotheses with Equality and Inequality Constraints:\\ A tutorial using the Bayes Factor via the Encompassing Prior Approach}
\shorttitle{The Bayes Factor via the Encompassing Prior Approach}

@@

\abstract{When conducting a study, researchers usually have expectations based on hypotheses or theoretical perspectives they want to evaluate. Equality and inequality constraints on the model parameters are used to formalize researchers' expectations or theoretical perspectives into the so-called “\textit{informative hypotheses}”. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are unsuitable for testing complex informative hypotheses. An alternative approach is to use the Bayes Factor. In particular, the Bayes Factor based on the “\textit{encompassing prior approach}” allows researchers to easily evaluate complex informative hypotheses. This paper provides a detailed introduction to the Bayes Factor with encompassing prior. First, all steps and elements involved in the formalization of  informative hypotheses and computation of the Bayes Factor are described. Next, we apply this method to a real case scenario considering the attachment theory. Precisely, we analyzed the relative role of maternal and paternal attachment on children's social-emotional development by comparing the different theoretical perspectives debated in the literature.

\vspace{2em}
\begin{center}
\textbf{Translational Abstract}
\end{center}
\vspace{.5em}

\noindent The evaluation of hypotheses is one of the main goals of empirical research. In fact, when conducting a study, researchers usually want to evaluate the plausibility of specific hypotheses according to the collected data. These hypotheses can be formalized in accordance with researchers' expectations or theoretical perspectives into the so-called “\textit{informative hypotheses}”. However, traditional statistical approaches, such as the Null Hypothesis Significance Testing (NHST) or the model comparison using information criteria (e.g., AIC and BIC), are not suitable for evaluating the plausibility of complex informative hypotheses. The Bayes Factor via the “\textit{encompassing prior approach}”, instead, allows researchers to easily evaluate the plausibility of complex informative hypotheses. This paper provides a detailed introduction to the Bayes Factor with encompassing prior. First, all steps and elements involved in the formalization of  informative hypotheses and computation of the Bayes Factor are described. Next, we apply this method to a real case scenario considering the attachment theory. Precisely, we analyzed the relative role of maternal and paternal attachment on children's social-emotional development by comparing the different theoretical perspectives debated in the literature. This approach is of particular interest in applied research as it allows researchers to formalize hypotheses with great flexibility and to directly evaluate which hypothesis is the most supported by the data.
}

\keywords{Bayes Factor, Encompassing Prior, Informative Hypothesis, Equality and Inequality Constraints, Attachment Theory}

\authornote{
\vspace{-1cm}
\begin{flushleft}\footnotesize
\addORCIDlink{Claudio Zandonella Callegher}{0000-0001-7721-6318}\\
\addORCIDlink{Tatiana Marci}{0000-0002-2813-0312}\\
\addORCIDlink{Pietro De Carli}{0000-0001-6538-2175}\\
\addORCIDlink{Gianmarco Altoè}{0000-0003-1154-9528}

Data used for the illustrative application are part of a larger project on the socioemotional development of school-age children, which was approved by the Ethics Committee of the School of Psychology at the University of Padova (protocol \#[1838-2016]). All materials and data are available at [TODO].

Correspondence concerning this article should be addressed to Claudio Zandonella Callegher, Department of Developmental Psychology and Socialisation, University of Padua, Via Venezia, 8 - 35131 Padova, Italy.  E-mail: claudiozandonella@gmail.com
\end{flushleft}
}

@@

\section{Introduction}

@@

The dominant statistical approach to evaluate research hypotheses in psychology is the Null Hypothesis Significance Testing (NHST). In the literature, however, the utility and validity of NHST are largely debated \parencite{wassersteinMovingWorld052019}. This approach presents indeed several limitations. First, the NHST places a narrow focus on statistical testing rather than on the formalization of hypotheses. This has usually led researchers to evaluate data against the catch-all null-hypothesis that nothing is going on rather than testing their specific expectation [and the alternative hypothesis is rarely formalized]. Second, the p-value does not quantify the evidence in favour of one hypothesis, thus, it is not possible to \textit{“accept”} a hypothesis but only to \textit{“reject”} it. This has some drawbacks since, in case of not significant results, the researchers are left in a state of indecision. Third, NHST does not allow testing multiple hypotheses at the same time. Using the NHST the null hypothesis is tested only against a single alternative. Fourth, NHST is unsuitable for testing broad classes of hypotheses with equality and inequality constraints \parencite{mulderSimpleBayesianTesting2019}. Expectations can be evaluated using one-side tests; however, when more groups or more variables are involved, it is not possible to evaluate complex parameter constraints that reflect researchers' expectations \parencite{vandeschootIntroductionBayesianModel2011}.

Model comparison is a different approach that allows researchers to compare multiple hypotheses and identify which is the most supported by the data \parencite{mcelreathStatisticalRethinkingBayesian2020}. Hypotheses are first formalized as statistical models according to researchers expectations or theoretical perspectives. Subsequently, it is possible to evaluate which is the most supported model among those considered according to the data. To do that a popular approach is to use information criteria such as the AIC or BIC criteria that estimate models ability to predict new data \parencite{wagenmakersAICModelSelection2004,akaike1973a, schwarzEstimatingDimensionModel1978}. Model comparison has several advantages and, in particular, it allows to directly compute the relative plausibility of each model given the data and the set of models considered. However, model comparison using information criteria is not appropriate to evaluate hypotheses that include complex parameter constraints reflecting researchers' expectations. In fact, information criteria evaluate models complexity according to the number of parameters, but they do not take into account possible order constraints on the parameters.

An alternative criterium to evaluate research hypotheses in a model comparison is the Bayes Factor. During the last 25 years, there has been an increasing interest in the Bayes Factor and its use has been proposed as the solution to the critical issues of the NHST \parencite{heckReviewApplicationsBayes2020, mulderEditorsIntroductionSpecial2016}. Although it has its own limitations and it does not solve the fundamental issues of the misuse of statistical techniques \parencite{schadWorkflowTechniquesRobust2021, gelmanBayesianDataAnalysis2013}, the Bayes Factor offers some clear advantages. In particular, the Bayes Factor allows us to compare hypotheses obtaining a relative index of evidence, like the information criteria, but it also allows us to easily compare complex research hypotheses. In fact, so-called \textit{“informative hypotheses”} (i.e., hypotheses containing information about the ordering of the model parameters) can be formalized according to researchers' expectations or theoretical perspectives using equality and inequality constraints \parencite{vandeschootIntroductionBayesianModel2011}. Subsequently, these hypotheses can be compared against each other using the Bayes Factor.

The evaluation of informative hypotheses is of particular interest to the researchers as it allows them to assess specific complex expectations and theoretical perspectives directly. In the literature, a particular approach allowing to easily compute the Bayes Factor with complex informative hypotheses has received increasing attention: \textit{“the Bayes Factor with encompassing prior”}. \textcite{vandeschootIntroductionBayesianModel2011} presented a general introduction to informative hypotheses testing using the Bayes Factor with the encompassing prior approach, whereas \textcite{hoijtinkInformativeHypothesesTheory2012} offered a more detailed description of the development of this method. Other studies, instead, considered the application of this approach with specific statistical models; for example: mixed effect models \parencite{katoBayesianApproachInequality2006}, evaluation of correlation coefficients \parencite{mulderBayesFactorsTesting2016} and multinomial models \parencite{heckMultinomialModelsLinear2019}. In addition, \textcite{guBayesianEvaluationInequality2014} proposed a general approximate procedure to evaluate inequality constraints in a wide range of statistical models.

However, all studies mentioned above considered only informative hypotheses with inequality constraints. Evaluating in the same hypothesis equality and inequality constraints, instead, it was not possible and equality constraints had to be approximated to “\textit{about equality constraints}” (i.e., a equality constraints of type $\theta_i = \theta_j$ is approximated to $|\theta_i - \theta_j| < \xi$ for a value of $\xi$ small enough).

@@

The development and implementation of this approach are of particular interest because, although with its limits, it allows researchers to directly evaluate their expectations and hypotheses. The available literature, however, is rather technical. The complexity of these articles makes it difficult for researchers not familiar with this approach to clearly understand all the steps involved. On the other hand, articles offering a general introduction to the Bayes Factor do not provide enough details to allow readers to autonomously apply this method to their problems, but they usually rely on ad-hoc solutions implemented in some statistical software \parencite[for example consider][]{hoijtinkTutorialTestingHypotheses2019a}.

@@

Prior specification is an important element as the resulting Bayes Factor value is affected by the prior choice. This is particularly relevant in the case of equality constraints \parencite{lindleyStatisticalParadox1957, bartlettCommentLindleyStatistical1957}, whereas inequality constraints are not affected as long as the prior is symmetric and centred to the focal point of interest (this feature will be further discussed in Section~\ref{sec:adj-prior}).

@@

Both, \textcite{guApproximatedAdjustedFractional2018, mulderSimpleBayesianTesting2019} based their approach on the fractional Bayes Factor. Starting from a non-informative prior, a minimal fraction of the data is used to obtain a posterior that is subsequently used as proper prior, we refer to it as \textit{fractional prior} (consider  Lindley's \citeyear{lindleyBayesianStatisticsReview1972} famous quote \textit{“Today’s posterior is tomorrow’s prior”}). The remaining part of the data is used to compute the Bayes Factor. The two approaches, however, have an important difference. \textcite{guApproximatedAdjustedFractional2018} approximated the obtained fractional prior (as well as the posterior, see Section~\ref{sec:posterior-encompassing}) to a (multivariate) normal distribution. In fact, due to large-sample theory \parencite{gelmanBayesianDataAnalysis2013}, parameter posterior distribution can be approximated to a (multivariate) normal distribution. On the contrary, \textcite{mulderSimpleBayesianTesting2019} provided an analytic solution in the case of linear regression models, obtaining an accurate quantification of the distribution.

All methods discussed above suggest objective procedures to avoid arbitrary prior specification. Nevertheless, using subjective (reasonable) priors according to previous information or experts' indications is still a possible approach. However, prior specification (even if obtained through an \textit{objective procedure}) affects the Bayes Factor results. Thus, it is fundamental to conduct a prior sensitivity analysis to evaluate the influence of prior specification on the results \parencite{schadWorkflowTechniquesRobust2021, duBayesFactorOnesample2019}.

@@

At this point, several authors emphasize the importance of centering the prior distribution on the constraints points of interest. Let's further discuss this issue in the next section.

@@

At this point, we have both prior and posterior distributions of the parameters of interests of the encompassing model. Before proceeding, however, let's underline some important aspects.

@@

Similarly, it is possible to compute the \textbf{marginal posterior density} $\pi_{adj}(\beta_E =0| Y, H_u)$ considering this time the estimated posterior mean vector $\bm{\hat{\beta}}$ and the estimated  posterior covariance matrix $\hat{\Sigma}_{\beta}$.

@@

To compute the Bayes Factor between two different informative hypotheses, we can simply consider the ratio of the Bayes Factor of each hypothesis against the encompassing model:
\begin{equation}
BF_{ij} = \frac{BF_{iu}}{BF_{ju}}
\end{equation}
In order to do that, however, it is necessary for the informative hypotheses considered to be comparable \parencite{guApproximatedAdjustedFractional2018}. That is, the intersection of the constrained region of each hypothesis must be non-empty. Thus, a set of informative hypotheses $\{H_1,\ldots, H_i\}$ is comparable if exist at least one solution to
\begin{equation}
\begin{bmatrix}
R_{1E}\\
R_{1I}
\end{bmatrix} \theta = \begin{bmatrix}
r_{1E}\\
r_{1I}
\end{bmatrix}, \ldots,\begin{bmatrix}
R_{iE}\\
R_{iI}
\end{bmatrix} \theta = \begin{bmatrix}
r_{iE}\\
r_{iI}
\end{bmatrix}.
\end{equation}

@@

The described approach using (multivariate) normal approximation can be straightforward applied when parameters of interest are unbounded. However, in the case of bounded parameters (i.e., variances, correlations, or probabilities), appropriate prior distributions are required to take into account the parameters bounded nature.

@@

These issues are relevant only when considering the prior distribution. In the case of the posterior distribution, instead, we can safely approximate it to a (multivariate) normal distribution thanks to the large-sample theory. Moreover, prior specification is relevant only for the parameters of interest (i.e., the parameters involved in the constraints) as nuisance parameters are integrated out in the Bayes Factor. Thus, the results will be insensitive to the actual choice of the prior for the nuisance parameters as long as it is vague enough.

@@

\section{A Case Study: Hypotheses Testing in Attachment Theory}

In this section, we propose a real case study evaluating different informative hypotheses within attachment theory. This will allow us to face the common issues found in real research applications. First, we provide the required background information regarding attachment theory and the study aims and characteristics. Subsequently, all the steps involved in the evaluation of informative hypotheses using the Bayes Factor with the encompassing prior approach are described. Finally, the results are briefly discussed considering the prior sensitivity analysis and limits of this approach.

@@

Attachment theory originates from the pioneering work of \textcite{bowlbyAttachmentLoss1969} and Ainsworth (\citeyear{ainsworthAttachmentExplorationSeparation1970}). They postulate that children in stressful situations actively seek proximity of the caregiver through some attachment behaviors (e.g., crying; moving towards the caregiver) in order to fulfil the evolutionary goal of protection from dangers. The main tenet of attachment theory is that the relationships with the caregivers that children develop in the early stages of their life (i.e., \textit{attachment bond}) will affect children’s social-emotional future development \parencite{cassidyHandbookAttachmentTheory2016}. Besides behavior, people construct mental representations, or working models, of the self and significant others based on their interpersonal experiences. Describing individual differences in self reported attachment mental representations, empirical research suggested a model based on two dimensional traits \parencite{mikulincerAttachmentAdulthoodStructure2007}: attachment avoidance (characterized by discomfort with closeness, excessive self-reliance, and difficulty in relying on others for obtaining security) and attachment anxiety (characterized by low self-worth and fear of abandonment and rejection). This perspective, initially developed with adult participants, has been extended to the study of self reported attachment representations also during childhood, even to middle childhood \parencite{kernsAttachmentMiddleChildhood2005, marciBriefExperiencesClose2019} . The two dimensions of attachment mental representations can be combined to provide four different attachment styles:

\begin{itemize}
  \item{\textbf{Secure Attachment}} - children who are securely attached display optimal emotional regulation and they consider the caregiver as a secure base.
  \item{\textbf{Anxious Attachment}} - anxious children manifest high levels of anxiety in stressful situations and their relationships with the caregivers is ambivalent displaying anger or helplessness.
  \item{\textbf{Avoidant Attachment}} - avoidant children mask distress in stressful situations displaying little emotions and their relationships with the caregivers is characterized by little involvement.
  \item{\textbf{Fearful Attachment}} - fearful children lack adequate emotional regulation in stressful situations with the risk of displaying non organized behaviors.
\end{itemize}

Attachment theory is one of the main and most supported theories in psychology \parencite{cassidyHandbookAttachmentTheory2016}. In the literature, however, there is still an open debate on the relative role of mother and father attachment on children's social-emotional development. Four different main theoretical perspectives have been identified \parencite{brethertonFathersAttachmentTheory2010}:
\begin{itemize}
  \item{\textbf{Monotropy Theory}} - only the principal attachment figure (usually the mother) has an impact on children's development.
  \item{\textbf{Hierarchy Theory}} - the principal attachment figure (usually the mother) has a greater impact on children development than the subsidiary attachment figure (usually the father).
  \item{\textbf{Independence Theory}} - all attachment figures are equally important but they affect the children's development differently.
  \item{\textbf{Integration Theory}} - to understand the impact on children's development it is necessary to consider all attachment relationships taken together.
\end{itemize}
Contrasting results have been reported by studies investigating which is the “\textit{correct}” theory. No study, however, has tried to properly evaluate the different theoretical perspectives by directly comparing the different hypotheses.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{The Present Study}

@@

Children's social-emotional development was measured using the Strength \& Difficulties Questionnaire \parencite[SDQ;][]{goodmanWhenUseBroader2010} completed by the teachers. Separate scores for externalizing and internalizing problems were obtained as sum of the questionnaire items. For the pourpose of this work, we considered only externalizing problems  The distribution of externalizing problems ($\text{Mean} = \Sexpr{my_round(mean(data_cluster$externalizing_sum), 2)};\ \text{SD} =  \Sexpr{my_round(sd(data_cluster$externalizing_sum), 2)};\ \text{Median} = \Sexpr{my_round(median(data_cluster$externalizing_sum), 1)}$) is presented in Figure~\ref{fig:plot-externalizing-dist}.
<<plot-externalizing-dist, cache=TRUE, fig.cap="Distribution of externalizing problems ($n_{subj} = 847$).">>=
plot_externalizing_dist()
@

Externalizing problems according to attachment styles are reported in Table~\ref{tab:table-cluster-ext}. More information about the sample, descriptive statistics, cluster analysis, and analysis of internalizing problems can be found in the Supplemental Material available online \url{https://claudiozandonella.github.io/Attachment/}. All materials and data are available at [TODO].

@@

When interpreting the results, it is important to take into account the limits of the Bayes Factor. First of all, the selected hypothesis is relative to the data and the set of hypotheses considered. That means that we should not interpret the selected hypothesis as the only “\textit{correct}” one. Maybe we did not consider some important aspects and new hypotheses may reveal themselves to be actually much better than the previous ones. For example, regarding the attachment results, taking into account the interaction between children's gender and attachment may be fundamental (is it possible that attachment towards the father plays a different role between girls and boys?). This limit, however, becomes an advantage as it forces the researchers to focus on the definition of the hypotheses: to obtain good answers first we need to ask the right questions.

@@

As emphasized by \textcite[p.229]{guApproximatedAdjustedFractional2018}, “this class of informative hypotheses covers a much broader range of scientific expectations than the class of standard null hypotheses. In addition, by testing competing informative hypotheses directly against each other a researcher obtains a direct answer as to which scientific theory is most supported by the data”.
